---
title: "Earthquakes vignettes"
author: "JC"
date: "16 juli 2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Earthquakes vignettes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The 'earthquakes' package

## Overview
The 'earthquakes' package is a package to clean, visualise and map NOAA earthquakes data. The functions to clean the data are 'eq_clean_data' and 'eq_location_clean' functions. Functions to visualise the data are 'geom_timeline' and 'geom_timeline_label' functions and to map are 'eq_map' and 'eq_create_label' functions. Below there are a short description and an example of these functions.
The usd packages are readr, dplyr, magrittr, tidyr, stringi, scales, ggplot2, grid and leaflet.


## Cleaning data
The 'eq_clean_data' function 
 - creates a 'DATE' variable from 'YEAR', 'MONTH' and 'DAY' columns, 
 - adjusts the longitude and longitude variable into numerical variables
 - creates a 'locations' variable from 'LOCATION_NAME' column
This function works together with the 'eq_location_clean' function that turn the 'COUNTRY' column into title case.


```{r cleaningLocation_NOAAData}
fileName <-  system.file("inst/extdata/earthquakes.tsv.gz")
file <- readr::read_delim(fileName, delim = "\t")
locations_cleaned <- eq_location_clean(data, countryCol)
locations_cleaned
```


```{r cleaning_NOAAData}
fileName <-  system.file("inst/extdata/earthquakes.tsv.gz")
file <- readr::read_delim(fileName, delim = "\t")
file_cleaned <- eq_clean_data(file)
file_cleaned
```



## Visualising data
The 'geom_timeline' function creates a timeline with the following elements:
 - eacht timeline represents a country
 - the earthquake events are displayed as circles
 - the size of the earthquake indicted the magnitude of the earthquakes
 - the color of the circles represents the number of deaths of an earthquake
The location of each earthquake is created by the 'geom_timeline_label' function.

```{r visualising_NOAAData, eval=FALSE, include=FALSE}
fileName <-  system.file("extdata/earthquakes.tsv.gz", package = "earthquakes")
file <- readr::read_delim(fileName, delim = "\t")
data <- eq_clean_data(file) %>%
  filter(COUNTRY == c("China", "Usa", "Japan"),
         DATE >= "1999-01-01",
         DATE <= "2012-12-31")
ggplot(data, aes(x = DATE, y = COUNTRY, size = richterScaleValue, fill = DEATHS)) +
  geom_timeline() +
  theme(legend.position = "bottom") +
  geom_timeline_label(aes(x=DATE, label = locations, n_maxVar = EQ_PRIMARY, n_max=10)) +
  labs(x = "Date", y = "Country", fill = "# deaths", size = "Richter scale value") 
```


## Mapping 
The 'eq_map' function maps the NOAA earthquakes data on a geographical map using the 'leaflet' package. The earthquakes are circles and the size of the circles is the magnitude of the earthquakes. When hovering over an earthquake on the map, a pop-up window displays information about the location, the magnitude and the number of deaths of the respective earthquake. The latter is done by the 'eq_create_label' function.

```{r mapping_NOAAData, eval=FALSE, include=FALSE}
map <- readr::read_delim("inst/extdata/earthquakes.tsv.gz", delim = "\t") %>% 
  eq_clean_data() %>% 
  dplyr::filter(COUNTRY == "Mexico" & lubridate::year(DATE) >= 2000) %>% 
  dplyr::mutate(popup_text = eq_create_label(.)) %>% 
  eq_map(annot_col = "popupText")
map 
```
